Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy

### 2020-01-20 manuscript/manuscript.Rmd
Suparna edits

---
title: "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Christopher Newport University
  email: suparna.chaudhry@cnu.edu
  url: http://www.suparnachaudhry.com/
  twitter: SuparnaChaudhry
- name: Andrew Heiss
  affiliation: Brigham Young University
  email: andrew_heiss@byu.edu
  url: https://www.andrewheiss.com/
  twitter: andrewheiss
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State crackdown on non-governmental organizations (NGOs), particularly through legal restrictions on funding and advocacy, has become increasingly pervasive across the globe. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? With reduced funding for NGOs working on contentious issues, and an absence of philanthropic culture in various developing countries, many NGOs rely on funds from private donors in Western countries. Existing research has predominantly focused on individual-level determinants of philanthropy. However, we argue that domestic political environments of NGO host countries can also influence preferences of foreign private donors, and that crackdown on NGOs serve as a heuristic to donors of organizational deservingness. Using a survey experiment, we explore how this crackdown changes donors' preferences based on the issue area and funding sources of the NGO. We find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We conclude by discussing the implications of our results on the sustainability of civil society organizations working in repressive countries. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of Political Science, Christopher Newport University; Andrew Heiss, PhD, Visiting Assistant Professor, Romney Institute of Public Service and Ethics, Brigham Young University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of Political Science, Christopher Newport University, Newport News, VA 23606, 757-594-7821. Contact: suparna.chaudhry@cnu.edu
word-count: "4,081"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

The 2011 Egyptian uprising presented an opportunity for domestic non-governmental organizations (NGOs) and civil society actors to transform Egypt into a democratic state. However, just as NGOs began to take advantage of new levels of civic freedoms, successive Egyptian governments passed laws restricting these groups from receiving foreign funding or engaging in political advocacy. Egypt is not alone in its non-violent crackdown on NGOs. India, Bolivia, Hungary, China, Russia, and others have imposed funding and programmatic restrictions on NGOs  [@Carothers:2015; @DPR:2015]. These laws have tangible consequences for NGOs worldwide. Official aid flows have decreased substantially in countries that have imposed restrictive anti-NGO laws [@ChaudhryHeiss:2018; @DupuyPrakash:2018], and large foundations and government agencies have both reduced their contributions to NGOs in repressive countries [@Brechenmacher:2017].

In response to this dramatic shift in the NGO funding landscape, organizations have increasingly turned to *foreign* individual donors to offset the loss of governmental and foundation funding [@BanksHulmeEdwards:2015]. This is because donations from *domestic* donors in repressive countries are unreliable and inconsistent. Citizens in the Global South often cannot support local NGOs, may lack a culture of philanthropic giving, or may channel funds to groups working on non-contentious issues such as health and education [@DPR:2015; @RPC:2016; @Brechenmacher:2017]. Even when donors in repressive countries prefer to donate to NGOs working in contentious areas, they may be deterred by dismal tax incentives, threats to take away business permits, or a general fear of retribution [@HudsonInstitute:2015]. Therefore, NGOs have increasingly sought out funds from *foreign* individual donors in countries with institutions more amenable to philanthropy. Further, given exchange rates and wealth differentials, it is easier for local NGOs to raise modest amounts of money abroad, rather than investing resources to raise similar sums locally [@RonKairePandya:2018].

Given increased crackdown on NGOs across the world, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate? We argue that domestic political environments of NGO host countries can influence preferences of foreign private donors and that legal crackdowns on NGOs serve as a heuristic to donors of organizational deservingness. We use a survey experiment to measure the effect of crackdowns on stated donor behavior and find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate substantially more to legally restricted NGOs. This effect persists when mediated by other structural heuristics like NGO issue area and its main funding sources.

@@

The decision to donate to an NGO is rarely rational. In theory, perfectly rational donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation. Instead, research has shown that donors rely on shortcuts, signals, and heuristics to determine the trustworthiness of NGOs. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. In addition to organizational characteristics that signal an NGO's trustworthiness and deservingness, structural factors—including the nationality of the targets of NGO programs and whether an organization works abroad—can also serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017].

We propose that the regulatory relationship between international NGOs and their host governments serves as a signal of NGO trustworthiness and deservingness. NGO legislation is not inherently restrictive, and governments frequently use regulations to benignly routinize the behavior of NGOs [@BloodgoodTB]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy. These stricter anti-NGO laws can signal to donors that NGOs undertake crucial work abroad, which is why governments perceive them as threatening and crack down on them. Therefore, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated:

@@

@fig:causal-path represents this relationship more formally. Legal crackdowns (**C**) have a direct effect on donor outcomes (**Y**; donation likelihood and amount) because of their heuristic function. Legal crackdowns do not occur in isolation, however. Countries are more likely to repress NGOs when organizational missions (**I**) are in tension in with government preferences and when organizations receive substantial funding (**F**) from abroad [@DPR:2015]. Each of these organizational characteristics increase the likelihood of crackdown but also simultaneously serve as donor heuristics and influence perceptions of NGO deservingness.

@@

The *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences—can influence donor behavior. Donors commonly support programs that have easily measurable goals and successes, and NGOs working on regime-compatible issues like health, education, and humanitarian relief rarely challenge the government [@Bush:2015; @Kiai:2017]. As such, their programmatic output is easily quantifiable and donors can see the results of their charity more readily. In contrast, NGOs working on more contentious human rights issues can pose political challenges to regimes. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. Due to the nature of the political and social issues they work on, programmatic effectiveness is also harder to measure and quantify [@Bush:2015]. Thus, donors may believe that it is more difficult to improve human rights in the target country and reduce their donations.

@@

The *source of NGO funding* can also influence donor behavior. Government grants to NGOs can often crowd out private funding [@AndreoniPayne:2003]. Donors may feel less inclined to contribute to NGOs receiving government funding, as they might perceive that their tax money already funds the organization. Government funding can also signal a lack of independence from the government. Organizations that receive substantial government funding frequently avoid programming that would question the government even if it were in line with their mission [@Stroup:2012]. In contrast, donors who see that organizations are privately funded may feel that they can also contribute and help.

As funding restrictions are one of the most common forms of anti-NGO regulations [@DPR:2015], organizational funding can influence anti-NGO crackdowns (see @fig:causal-path). As such, the combination of legal crackdown and funding may reshape donor perceptions of deservingness. Private donors may be more likely to punish government-funded NGOs that face restrictions. This is because legal trouble with the group's main funder could fuel private donor mistrust. Donors thus may be more likely to donate more to NGOs that rely on private funding:

@@

@@

Emphasizing legal difficulties increased donations to government-funded humanitarian NGOs by `r percent_0(filter(amount, title == "Humanitarian assistance | Government")$pct_change)`, or `r md_dollar(abs(filter(amount, title == "Humanitarian assistance | Government")$mu_diff))` (`r filter(amount, title == "Humanitarian assistance | Government")$p_neq0`). The crackdown condition causes this increase despite some hesitation from respondents who often indicated that government-funded NGOs do not deserve private donations. Respondents reasoned since IRC "is already receiving funding from governments, so funding from private individuals doesn't seem as necessary. I'd prefer to give my money to an organization that is primarily run by private donations"[^6486084] or because "[m]y tax paying dollars go to the government and if the government is helping to fund it then they are in turn getting some of my money."[^9509936] This hesitation is not universal, however, and in aggregate most respondents chose to give substantially more. Many echoed this respondent, who identified the crackdown condition as one of the reasons for giving the full hypothetical \$100: "I believe the work of such organizations has an urgency that affects many refugees and their families. If they can function without being affected by the harsh laws against non-profits, I would want them to have as much help as possible to do their humanitarian work."[^1097165]

@@

# Conclusion

Dozens of countries have legally restricted NGO funding and programming. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many international NGOs have turned to foreign private donors as an additional funding source. Do ongoing legal crackdowns influence these donors' decision to donate? Our results show that legal restrictions provide donors with an additional signal of NGO deservingness and efficiency and can change both the likelihood that donors will give and the amount they will give. While crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate `r percent_0(filter(amount, title == "Crackdown")$pct_change)` more to legally restricted NGOs. This effect persists when mediated by other heuristics, with respondents giving `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` more to humanitarian NGOs, `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)` more to privately-funded human rights NGOs, and `r percent_0(filter(amount, title == "Humanitarian assistance | Government")$pct_change)` more to government-funded humanitarian NGOs. Anti-NGO crackdowns can thus serve as a signal of NGO effectiveness.

This study makes two important contributions. While recent work has looked at how state repression of NGOs affects official aid [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how repression affects private aid. Further, prior literature on the determinants of donation to NGOs focuses predominantly on individual-level factors. Our study measures important structural determinants that influence philanthropy and shows how these factors interact with organizational attributes to signal NGO deservingness. Additional research should determine the extent to which these preferences translate into behavior—here, we have only asked respondents to make a hypothetical donation—as well as test the generalizability of our findings to other issue areas.

This has several implications for policy and future research. NGOs may benefit from publicizing when they are targets of government crackdown. While this may not be enough to sway the average U.S. citizen, our results show that communicating this information can convince already-likely donors to donate more to besieged groups. Our study also highlights the importance of framing NGO operations. Previous research on human rights micro-mobilization has shown that the way people process information can affect their increased willingness to help others in need [@Slovic:2007]. Our results show that donors are likely to view legally besieged human rights NGOs with suspicion. However, human rights groups undertake a variety of programs, and communicating the nature of their work and specifying the main target of the organization's efforts can allay the public's fears. Finally, if NGOs signal to donors that receive a majority of their funding from non-governmental sources, they may be able to convince average donors that their contributions might help the NGO resist the crackdown. Moreover, since private funders typically tend to be less interested in the domestic politics of recipient countries compared to governmental funders, relying on private donations is less likely to attract the ire of authoritarian governments fearful of foreign intervention.

@@

>>>>

---
title: "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Christopher Newport University
  email: suparna.chaudhry@cnu.edu
  url: http://www.suparnachaudhry.com/
  twitter: SuparnaChaudhry
- name: Andrew Heiss
  affiliation: Brigham Young University
  email: andrew_heiss@byu.edu
  url: https://www.andrewheiss.com/
  twitter: andrewheiss
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State crackdown on non-governmental organizations (NGOs), particularly through legal restrictions on funding and advocacy, has become increasingly pervasive across the globe. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? With reduced funding for NGOs working on contentious issues, and an absence of philanthropic culture in various developing countries, many NGOs rely on funds from private donors in Western countries. Existing research has predominantly focused on individual-level determinants of philanthropy. However, we argue that structural factors such as domestic political environments of NGO host countries can also influence preferences of foreign private donors, and that crackdown on NGOs serve as a heuristic to donors of organizational deservingness. Using a survey experiment, we explore how this crackdown changes donors' preferences based on the issue area and funding sources of the NGO. We find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We conclude by discussing the implications of our results on the sustainability of civil society organizations working in repressive countries. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of Political Science, Christopher Newport University; Andrew Heiss, PhD, Visiting Assistant Professor, Romney Institute of Public Service and Ethics, Brigham Young University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of Political Science, Christopher Newport University, Newport News, VA 23606, 757-594-7821. Contact: suparna.chaudhry@cnu.edu
word-count: "4,081"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

Following the 2011 Egyptian uprising, just as domestic non-governmental organizations (NGOs) began to take advantage of new levels of civic freedoms, successive Egyptian governments passed laws restricting these groups from receiving foreign funding or engaging in political advocacy. Egypt is not alone in its crackdown on NGOs. India, Bolivia, Hungary, Russia, and others have imposed funding restrictions on NGOs  [@Carothers:2015; @DPR:2015]. Consequently, official aid flows as well as grants from large foundations have considerably decreased to NGOs in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018].

In response to this dramatic shift in the NGO funding landscape, organizations have increasingly turned to *foreign* individual donors to offset the loss of governmental and foundation funding [@BanksHulmeEdwards:2015]. This is because donations from *domestic* donors in repressive countries are unreliable and inconsistent. Citizens in the Global South often cannot support local NGOs, may channel funds to groups working on non-contentious issues such as health and education [@DPR:2015; @RPC:2016; @Brechenmacher:2017], or may be deterred by dismal tax incentives, threats to take away business permits, or a general fear of retribution [@HudsonInstitute:2015]. Therefore, NGOs have increasingly sought out funds from *foreign* individual donors in countries with institutions more amenable to philanthropy. By 2016, the amount directed towards civil society groups working in international affairs increased to \$22.03 billion [@givingusa:2017]. As Daniel Runde pointed out in a statement before a U.S. Senate subcommittee on multilateral institutions, “U.S. philanthropy is significant [for global philanthropic efforts] but it is mainly Ms. Smith and Mr. Johnson, [n]ot Mr. and Mrs. Gates” [@Runde:2017].

Given increased crackdown on NGOs across the world, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate? Given information asymmetry and time constraints, individual donors often use heurestics to simply their decision-making [@TverskyKahneman:1974]. While much of this research has focused on heuristics from social networks and organizational practices [@CrosonShang:2011; @Tremblay-BoirePrakash:2017], in this research note, we argue that domestic political environments of NGO host countries can also serve as one such structural heuristic that can influence individual donor decisions. In particular, legal crackdowns on NGOs serve as a heuristic to donors of organizational deservingness and when donors are provided salient information about crackdown on NGOs and its impact on NGO activities, it can influence donor preferences and behavior. We use a survey experiment to measure the effect of crackdowns on stated donor behavior and find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate substantially more to legally restricted NGOs. This effect persists when mediated by other organizational-level heuristics like NGO issue area and its main funding sources.

This research note two important contributions. While recent work has looked at how state repression of NGOs affects official aid [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how repression affects private aid. Further, while there is a robust existing literature on the individual-level determinants of philanthropy, far less is known about structural determinants that influence philanthropy, particularly towards international causes, and how these factors interact with organizational attributes to signal NGO deservingness.

Below, we first summarize existing research on the determinants of philanthropic-giving to NGOs and lay out our expectations regarding how crackdowns on NGOs may affect charitable-giving. We then describe our survey and present the results. We conclude with implications for NGOs working in repressive countries and lay out questions ripe for future research.

@@

Substantial research on charitable-giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011], and more recently, the effect of social networks on philanthropy [@HerzogYang:2018]. However, lesser research on focused on the impact of structural factors on the choice to donate. Moreover, research on individual donor behavior has overwhelming looked at giving to local, rather than international causes. While most research on giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014], we explore individual giving to such causes becauses on individuals may rely on signals or heurestics that differ from the benchmarks that large donors (whether official or private) rely on.

The decision to donate to an NGO is rarely rational. In theory, perfectly rational donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Existing research shows that information on the efficiency of NGOs merely steers rather than encourage or discourage overall donations [@ryazanov:efficiency]. Similarly, the effects of changes in charity watchdog rations and nonprofit accountability systems on individual donors is also mixed - showing that donors do not have the time to research charities thoroughly [@Bekkers:2010; @SzperPrakash:2011]. Instead, donors respond to a host of heurestics and signals when deciding whether to donate to an NGOs, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009].

We argue that in addition to the above factors that signal an NGO's trustworthiness and deservingness, structural factors—including the nationality of the targets of NGO programs and whether an organization works abroad—can also serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017]. We propose that the regulatory relationship between international NGOs and their host governments serves as a signal of NGO trustworthiness and deservingness. NGO legislation is not inherently restrictive, and governments frequently use regulations to benignly routinize the behavior of NGOs [@BloodgoodTB]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy. These stricter anti-NGO laws can signal to donors that governments perceive these groups as threatening and may crack down on them.

We argue that donors may process legal crackdowns differently from organizational-level heuristics - such as NGO missions, visions, values- because donors may be able to influence these latter attributes, at least to some extent. On the other hand, donors may perceive that they can do little to change the domestic political environment of another country. Such crackdowns can signal to donors that the government may eventually target other NGOs as well, which can increase donor urgency to support legally beseiged NGOs, even if donors are not completely satisfied with some organizational-level features of a group. As such, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated:

@@

@fig:causal-path represents this relationship more formally. Legal crackdowns (**C**) have a direct effect on donor outcomes (**Y**; donation likelihood and amount) because of their heuristic function. Legal crackdowns do not occur in isolation, however. Countries are more likely to repress NGOs when organizational missions (**I**) are in tension in with government preferences and when organizations receive substantial funding (**F**) from abroad [@DPR:2015]. Both of these organizational characteristics increase the likelihood of crackdown but also simultaneously serve as donor heuristics and influence perceptions of NGO deservingness.

@@

Another heurestic that individual donors may use for determining NGO deservingness is the *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences.[^targeting] Organizations that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. NGOs working on these issues can also frame their activities as in need of greater support. Perception of need, whether through personal experiences or media framing strongly influence individual decisions to donate to relief efforts [@BekkersWiepking:2010, 6-8].

In contrast, donors may shy from donating to groups working on political causes.  NGOs working on more contentious human rights issues can be challenging to regimes. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. Thus, donors may believe that it is more difficult to improve human rights in the target country and reduce their donations.

@@

The final heurestic that can influence individual donors that we examine is the *source of NGO funding*. Government grants to NGOs can often crowd out private funding for nonprofit organizations[@AndreoniPayne:2003], though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991]. Donors may feel less inclined to contribute to NGOs receiving home government funds because such funding can signal a lack of independence from the government. Organizations that receive substantial government funding frequently avoid programming that would question the donor government even if it were in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help.

As funding restrictions are one of the most common forms of anti-NGO regulations [@DPR:2015], organizational funding can influence anti-NGO crackdowns (see @fig:causal-path). As such, the combination of legal crackdown and funding may reshape donor perceptions of deservingness. Private donors may be less likely to donate to government-funded NGOs that face restrictions. This is because legal trouble with the group's main funder could fuel private donor mistrust. Donors thus may be more likely to donate more to NGOs that rely on private funding:

@@

In these three sets of hypotheses, we do not assume perfect knowledge among individual donors. Donors may not always know about anti-NGO laws in target countries, or who funds an NGO, when they make their contributions. Rather, the three heuristics we explore here are potential signals to donors and our intention is to vary these three factors to test how powerful these factors are in comparison to other individual and organizational-level attributes. This can ultimately help NGOs optimize their fundraising strategies and tailor their fundraising attempts to private individuals.

@@

Emphasizing legal difficulties increased donations to government-funded humanitarian NGOs by `r percent_0(filter(amount, title == "Humanitarian assistance | Government")$pct_change)`, or `r md_dollar(abs(filter(amount, title == "Humanitarian assistance | Government")$mu_diff))` (`r filter(amount, title == "Humanitarian assistance | Government")$p_neq0`). The crackdown condition causes this increase despite some hesitation from respondents who often indicated that government-funded NGOs do not deserve private donations. Respondents reasoned since IRC "is already receiving funding from governments, so funding from private individuals doesn't seem as necessary. I'd prefer to give my money to an organization that is primarily run by private donations"[^6486084] or because "[m]y tax paying dollars go to the government and if the government is helping to fund it then they are in turn getting some of my money."[^9509936] This hesitation is not universal, however, and in aggregate most respondents chose to give substantially more. Many echoed this respondent, who identified the crackdown condition as one of the reasons for giving the full hypothetical \$100: "I believe the work of such organizations has an urgency that affects many refugees and their families. If they can function without being affected by the harsh laws against non-profits, I would want them to have as much help as possible to do their humanitarian work."[^1097165] Thus, legal crackdowns can increase individual donor urgency to support an organization, and this heuristic is potentially processed differently compared to other organization-level heuristics or information from social networks.

@@

# Discussion: Does NGO crackdown affect individual philanthropy?

Dozens of countries have sought to restrict NGO funding and programming. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many international NGOs have turned to foreign private donors as an additional funding source. Since seeking out perfect information about a nonprofit and its activities, especially internationally, is costly and time-consuming, donors use heuristics when deciding whom to support. Existing research has shown that donors take cues and heuristis from social netowrks, but can other information cues can influence donor preferences? This research note argues that domestic political environments of NGO host countries can serve as one such heuristic. Our results show that crackdowns provide donors with an additional signal of NGO deservingness and efficiency and can change both the likelihood that donors will give and the amount they will give. We find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate `r percent_0(filter(amount, title == "Crackdown")$pct_change)` more to legally restricted NGOs. This effect persists when mediated by other heuristics, with respondents giving `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` more to humanitarian NGOs, and `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)` more to privately-funded human rights NGOs. Anti-NGO crackdowns can thus serve as a signal of NGO effectiveness.

This research note should be viewed as an exploratory study, and it raises numerous quesitions for future research. These include determining the extent to which these preferences translate into behavior—here, we have only asked respondents to make a hypothetical donation, but preferences may not reflect actual behavior. Future research can also test the generalizability of our findings to other issue areas.

These results have several implications for policy. NGOs may benefit from publicizing when they are targets of government crackdown. While this may not be enough to sway the average U.S. citizen, our results show that communicating this information can convince already-likely donors to donate more to besieged groups. Further, if NGOs signal to donors that receive a majority of their funding from non-governmental sources, they may be able to convince average donors that their contributions might help the NGO resist the crackdown. Moreover, since private funders typically tend to be less interested in the domestic politics of recipient countries compared to governmental funders, relying on private donations is less likely to attract the ire of authoritarian governments fearful of foreign intervention.

THIS PART ONLY IF WORDS LEFT: Our study also highlights the importance of framing NGO operations. Previous research on human rights micro-mobilization has shown that the way people process information can affect their increased willingness to help others in need [@Slovic:2007]. Our results show that donors are likely to view legally besieged human rights NGOs with suspicion. However, human rights groups undertake a variety of programs, and communicating the nature of their work and specifying the main target of the organization's efforts can allay the public's fears.

@@

[^targeting]: While anti-NGO legislation (in the form of barriers to entry, funding, and advocacy) can theoretically tatget all NGOs (or a large subset of them) within a state's borders, in practices, these laws are enforced more selectively against contentious groups that pose threats to states. See Chaudhry (2016) and Heiss (2017).
